
---
title: "Data Wrangling with Base R"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
   

---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```


```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
```

```{r, echo=FALSE}
xaringanExtra::use_webcam()
```

```{r xaringan-fit-screen, echo=FALSE}
xaringanExtra::use_fit_screen()
```


class: inverse, middle, center
# Data Frame

---
# What's a data frame?

- A data frame is a **table** of data 

- Each **column** contains values of one variable 

- Each **row** contains one set of values from each column.

--

- Example

| Name    | Salary | Start_Date | Gender |
|---------|--------|------------|--------|
| Rick    | 623.30 | 2012-01-01 | Male   |
| Dan     | 515.20 | 2013-09-23 | Male   |
| Jessica | 611.00 | 2014-11-15 | Female |
| Julia   | 729.00 | 2014-05-11 | Female |

---
# Creating a Data Frame

```{r}

df = data.frame(Name = c('Tom','Jack','Steve','Ricky'),
                Age = c(28,34,29,42),
                Rank = c(0,1,2,3))
        
df

```

---
class: middle, center

Import a data frame using `read.csv` 

Use `str` function to give an overview 

Correct data type using `as.factor`, `as.numeric`, `as.Date`, `as.character`

Check missing data with `sum(is.na())`, `colSums((is.na))`

---
# Read (Import) a Data Frame using `read.csv`

- Method 1: Download the file to your computer and read it. Usually the code and the data are in the same folder. 

```{r, eval=FALSE}
setwd("C:/Users/student/Documents/data2/Test")
df <-  read.csv('us-states.csv')
```

---
# Check the variables (columns) of the data

```{r}
names(df)
```

---
# Quick summary of the data

```{r}
str(df)
# The data has 5 columns for each variable with 51,078 rows for each observation
```

---
# Quick view of the data

```{r}
head(df)
```

---
# Check missing values

```{r}

sum(is.na(df))
colSums(is.na(df))
colMeans(is.na(df))
```

---
# Change the name of a variable

```{r}
names(df) <- c('Date', 'State', 'Fips', 'Cases','Deaths')
```

---
# Change the type of a variable

```{r}
# Check type
class(df$Date)
str(df)

# Change type
df$Date = as.Date(df$Date)

```

---
class: inverse, middle, center
# Simple Data Exploration

---
# Statistics of continuous/date variable

```{r}
mean(df$Cases, na.rm=TRUE)

max(df$Cases, na.rm=TRUE)
# Other operation on one variable
summary(df$Date)
```

---
# Frequency of Categorical variable
```{r}
length(table(df$State))
prop.table(table(df$State))
```
```{r}
weekdays = weekdays(df$Date)
```
```{r}
df$death2 = ifelse(df$Deaths==0,'no_death','has_death')
summary(df$Deaths)

library(dplyr)
df$death3 <- case_when(df$Deaths<838 ~ 'low_death', df$Deaths<12674 ~ 'mid_death',
                              df$Deaths>12674~'high_death')
```

```{r}
df[df$State=='RI',]
df2 <- df[(df$Date>="2021-01-01")&(df$Date<"2022-01-01")&(df$State=='RI'),]
mean(df$Death)
median(df$Death)
```

```{r}

```

---
# Correlation of two continuous variables

```{r}
# Correlation 

cor(df$Cumulative_cases, df$Cumulative_deaths)
```

---
# Calculate a continuous by a categorical

```{r}
by(df$New_deaths, df$WHO_region, mean)
```

---
# Calculate a continuous by a categorical

```{r}
by(data = df$New_deaths, INDICES = df$WHO_region, FUN = mean)
```

---
class: inverse, middle, center

# Create new variables

*It is common that to create new variables to analyze the data better*

---
# Binning Continuous variable to a categorical variable

Create the categorical variable `New_cases2` variable taking the values as follows

- `No_new_cases` if there is a no new cases that day
-  `Has_new_cases` if there is at least a new case that day

---
```{r}
df$New_cases2 <- ifelse(df$New_cases==0, 
                        'No_new_cases','Has_new_cases') 
```

---
```{r}
library(dplyr)
df$New_cases2 <- case_when(df$New_cases==0 ~ 'No_new_cases', 
                            TRUE ~ 'Has_new_cases') 
```


---
# Binning Continuous variable

Create the categorical variable `New_deaths2` variable taking the values as follows

- `low_death` if the number of `New_deaths` smaller 2

- `mid_death` if the number of `New_deaths` from 2 to 5

- `high_death` if the number of `New_deaths` greater than 5

---
```{r}
df$New_deaths2 <- case_when(df$New_deaths<2 ~ 'low_death',
                             df$New_deaths<5 ~ 'mid_death',
                             TRUE~'high_death')
table(df$New_deaths2)
```

---
# Binning categories of a categorical variable

- If a variable has too many categories, we may want to group those categories to fewer categories

```{r}

# Create a variable month
library(lubridate)
df$month <- month(df$Date_report, label = TRUE)


# group months into fewer categories
df$month2 <- case_when(df$month %in% c('Sep','Oct','Nov','Dec') ~ 'fall_semester',
                       df$month %in% c('Feb','Mar','Apr','May') ~ 'spring_semester',
                       TRUE~'break')
  
```

---
class: inverse, middle, center

# Subsetting

---
# Subsetting by columns

```{r}
# by columns
df1 <- df[,c(1,3)]
head(df1)
```

---
# Subsetting by columns

```{r}
# by columns
df1 <- df[, c('Date_report','New_cases')]
head(df1)
```

---
# Subsetting by rows

```{r}
# by rows
df1 <- df[c(1:5),]
head(df1)
```

---
# Subsetting by columns and rows

```{r}
# by rows
df1 <- df[c(1:5), c('Date_report','New_cases')]
head(df1)
```



---
# Subsetting by condition

```{r}
df1 <- df[df$New_deaths>10,]
df2 <- df[df$Country=='United States of America',]
df3 <- df[df$Date_report>='2021-01-01',]

df4 <- df[(df$Country=='United States of America')&
            (df$Date_report>='2021-01-01'),]

```

---
# Logical Operators

| Operator  	| Description              	|
|-----------	|--------------------------	|
| <         	| less than                	|
| <=        	| less than or equal to    	|
| >         	| greater than             	|
| >=        	| greater than or equal to 	|
| ==        	| exactly equal to         	|
| !=        	| not equal to             	|
| !x        	| Not x                    	|
| x &#124; y    	| x OR y                   	|
| x & y     	| x AND y                  	|
| isTRUE(x) 	| test if X is TRUE        	|

---
class: inverse, middle, center
# Some Examples

---
class: middle, center
Find the average new cases by region in 2021. 

---
```{r}
df1 <- df[df$Date_report>='2021-01-01',]
by(df1$New_cases, df1$WHO_region, mean)
```

---
class: middle, center
Find the median new deaths by weekday in the US

---
```{r}
df$weekdays <- weekdays(df$Date_report)
df1 = df[df$Country=='United States of America',]
by(data = df1$New_deaths, INDICES = df1$weekdays, FUN = median)
```

