
---
output: # rmarkdown::github_document
  html_document:
    theme: cerulean
    toc: true
  pdf_document: default
  word_document: default
title: "Assignment 5: Part 1"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment5_part1.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

**The data:** 

This assignment works with the Adult Census Data that can be downloaded at [this link.](https://bryantstats.github.io/math421/data/adult_census_missing.csv)

-------

**Questions**

1. Use `read_csv` to import the data.  Show the number of NAs for each columns.
```{r}
library(tidyverse)
df = read_csv('adult_census_missing.csv')
```


2. Using function `aggr` the VIM package to plot the number of NAs for each column. 
```{r}
library(VIM)
aggr(df)
```


3. Find other forms of missing values. Hint: You can use the `table` function to check if there are suspicious categories (Unknown, for example) in a variable. What all other forms of missing values shown in the data. 
```{r}
df <- replace(df, df == 'Unknown' |
                  df == 'Not Known' |
            df == '?' , NA)
aggr(df)
df2 = drop_na(df)
view(df2)
"29,865"
```
```{r}
df3 = df %>% fill(workclass, education, marital.status,occupation,relationship,native.country,sex, .direction = 'updown')

colSums(is.na(df3))

```


```{r}

df4 = df

median_age <- median(df4$age, na.rm=TRUE)
df4$age <- replace_na(df4$age, median_age)
median_capital.gain <- median(df4$capital.gain, na.rm=TRUE)

df4$capital.gain <- replace_na(df4$capital.gain, median_capital.gain)
majority_workclass <- names(which.max(table(df$workclass)))
df4$workclass <- replace_na(df4$workclass, majority_workclass)
majority_education <- names(which.max(table(df$education)))
df4$education <- replace_na(df4$education, majority_education)
majority_marital.status <- names(which.max(table(df4$marital.status)))
df4$marital.status <- replace_na(df4$marital.status, majority_marital.status)
majority_occupation <- names(which.max(table(df4$occupation)))
df4$occupation <- replace_na(df4$occupation, majority_occupation)
majority_relationship <- names(which.max(table(df$relationship)))
df4$relationship <- replace_na(df4$relationship, majority_relationship)
majority_sex <- names(which.max(table(df4$sex)))
df4$sex <- replace_na(df4$sex, majority_sex)
majority_capital.gain <- names(which.max(table(df4$capital.gain)))
df4$capital.gain <- replace_na(df4$capital.gain, majority_capital.gain)
majority_native.country <- names(which.max(table(df4$native.country)))
df4$native.country <- replace_na(df4$native.country, majority_native.country)

colSums(is.na(df4))
```






